The Effects of weather on maize yields: New evidence from Kenya

### 2018-12-07 writing/draft3/draft9.tex
working hard
writing editing

%Uncomment next line if AMS fonts required
%\usepackage{iopams}  
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage{color}
\usepackage{pdfpages}
\usepackage{float}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{amsbsy}
\usepackage[none]{hyphenat}
\usepackage{footnote}
\usepackage{tabularx}
\usepackage{epstopdf}
\usepackage{setspace} % for double spacing, maybe remove later
\usepackage{mathrsfs} 
\usepackage[hyphens]{url}
\usepackage[english]{babel}
\usepackage{harvard}

@@

@@

	There are two predominant rainfall regimes in Kenya. The Arid and semi-arid (ASAL) counties mostly exhibit bi-modal precipitation patterns with long rains lasting from March to May and short rains occurring between October and December. In the non-ASAL counties, the single rainy season usually starts in March and lasts until August. Following the precipitation patterns and closely related planting and harvesting calendar, we computed yearly values of various weather characteristics as follows: For the ASAL counties, we used daily data covering October, November and December of the previous year and March, April, May of the current year and for the non-ASAL counties we used daily data covering March to August of the current year.

@@

To find the most suitable set of predictors, we adopted a step-down model building approach using the Satterthwaite's method to determine the p-values of the individual t-tests \cite{lmerTest}. That is, we started with a model with all possible weather measures in both fixed effects and random effects and we consequently removed the insignificant variables until we left with a model with only significant predictors in fixed and random effects. For both fixed and random effects we considered the level of significance $\alpha=0.05$.

According to the conditional Lagrange multiplier (LM) test for serial correlation in the idiosyncratic component of the errors (developed by Baltagi \& Li 1991 and Baltagi \& Li 1995), the errors do exhibit a within group correlation structure (the p-value of the LM test is $9.48\times10^{-14}$). To investigate the autocorrelation structure of errors, we estimated all variants of the preferred model with autocorrelation structure of a form ARMA($p$,$q$) such that $p\leq2$ and $q\geq2$. Comparing the AIC criteria and using the corresponding likelihood ratio statistics, we found out that the most appropriate correlation structure of the errors is ARMA($1$,$1$). The AIC criterion see table~\ref{ARMA} in the appendix..

@@

\begin{table}[H]
\caption{\label{ASAL}\textbf{List of Arid and semi-arid (ASAL) and non-ASAL counties}}

@@

\begin{indented}
\item[]\begin{tabular}{@{}llll}
\\[-1em]
\br
\\[-1em]
\textbf{Error autocorrelation structure}&\textbf{AIC}&\textbf{Likelihood ratio}&\textbf{p-value}\\
\mr
\textit{None}& $2205.1$& & \\
ARMA($1$,$0$)& $2129.2$& $77.9$&$<0.0001$ \\
ARMA($0$,$1$)&  $2144.9$& & \\
\textbf{ARMA($\boldsymbol{1}$,$\boldsymbol{1}$)}& $\boldsymbol{2122.2}$& $\boldsymbol{24.7}$&$\boldsymbol{<0.0001}$ \\
ARMA($2$,$1$)& $2124.0$& $0.1$&$0.7066$ \\
ARMA($1$,$2$)&  $2124.2$& & \\
ARMA($2$,$2$)&$2125.8$& $0.3$&$0.5681$ \\
ARMA($3$,$2$)& $2128.0$& $0.1$&$0.7312$ \\
ARMA($2$,$3$)&  $2127.6$& & \\
ARMA($3$,$3$)&\multicolumn{3}{c}{\textit{Failed to converge}} \\
\br
\end{tabular}
\end{indented}
\end{table}

>>>>

%Uncomment next line if AMS fonts required
%\usepackage{iopams}  
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage{color, colortbl}
\usepackage{pdfpages}
\usepackage{float}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{amsbsy}
\usepackage[none]{hyphenat}
\usepackage{footnote}
\usepackage{tabularx}
\usepackage{epstopdf}
\usepackage{setspace} % for double spacing, maybe remove later
\usepackage{mathrsfs} 
\usepackage[hyphens]{url}
\usepackage[english]{babel}
\usepackage{harvard}

@@

\definecolor{Gray}{gray}{0.9}

@@

	There are two predominant rainfall regimes in Kenya. The arid and semi-arid (ASAL) counties exhibit mostly bi-modal precipitation patterns with long rains lasting from March to May and short rains occurring between October and December. In the non-ASAL counties, the single rainy season usually starts in March and lasts until August. Following the precipitation patterns and closely related planting and harvesting calendar, we computed yearly values of various weather characteristics as follows: For the ASAL counties, we used daily data covering October, November and December of the previous year and March, April, May of the current year and for the non-ASAL counties we used daily data covering March to August of the current year.

@@

To find the most suitable set of predictors, we adapted a step-down model building approach using the Satterthwaite's method to determine the p-values of the individual t-tests \cite{lmerTest}. To be more specific, we started with a model which included the complete set of our weather measures in both fixed effects and random effects. In the following steps, we were sequentially removing the insignificant variables until we were left with a model with only significant predictors in fixed and random effects. For both fixed and random effects we considered the level of significance $\alpha=0.05$.

According to the conditional Lagrange multiplier (LM) test for serial correlation in the idiosyncratic component of the errors (developed by Baltagi \& Li 1991 and Baltagi \& Li 1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $9.48\times10^{-14}$). To investigate the autocorrelation structure of the errors, we estimated several variants of our preferred specification with the autocorrelation structure of a form ARMA($p$,$q$). In particular, we estimated all varieties of our model with ARMA($p$,$q$) errors such that $p\leq2$ and $q\leq2$. Comparing the AIC criteria and using the corresponding likelihood ratio statistics, we found that the most appropriate correlation structure of the errors was ARMA($1$,$1$). The AIC criterion of the model with ARMA($1$,$1$) error structure was $2122.2$ while the AIC criterion of the model without any correlation structure was $2205.1$ and the AIC criterion of the model with ARMA($1$,$0$) errors was $2129.2$. The likelihood ratio test of comparison of

@@

\begin{table}[H]
\caption{\label{ASAL}\textbf{List of arid and semi-arid (ASAL) and non-ASAL counties}}

@@

\begin{indented}
\lineup
\item[]\begin{tabular}{@{}llll}
\\[-1em]
\br
\\[-1em]
\textbf{Error autocorrelation structure}&\textbf{AIC}&\textbf{Likelihood ratio}&\multicolumn{1}{c}{\textbf{p-value}}\\
\mr
\textit{None}& $2205.1$& & \\
ARMA($1$,$0$)& $2129.2$& $77.9$&$<1\times10^{-4}$ \\
ARMA($0$,$1$)&  $2144.9$& & \\
\rowcolor{Gray}ARMA(${1}$,${1}$)& ${2122.2}$&${24.7}$&${<1\times10^{-4}}$ \\
ARMA($2$,$1$)& $2124.0$&$\00.1$&$\0\00.7066$ \\
ARMA($1$,$2$)&  $2124.2$& & \\
ARMA($2$,$2$)&$2125.8$& $\00.3$&$\0\00.5681$ \\
ARMA($3$,$2$)& $2128.0$& $\00.1$&$\0\00.7312$ \\
ARMA($2$,$3$)&  $2127.6$& & \\
ARMA($3$,$3$)&\multicolumn{3}{c}{\textit{Failed to converge}} \\
\br
\end{tabular}
\end{indented}
\end{table}
